---
title: "BayesMultiMode"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(dpi=600)
```

This package provides functions for estimating the number of modes and their associated locations in mixtures of discrete distributions using Bayesian methods. The testing approach works in three stages. First, a mixture distribution is estimated on the data using MCMC methods. The number of mixture components is allowed to be unknown a priori and estimated using a Spare Finite Mixture (SFM) algorithm. Second, a simple post-processing is applied to the MCMC output where a given number of iterations are used for burn-in and empty components are discarded. Third, post-processed MCMC draws are used to compute the number of modes as well as their locations and posterior probabilities. 

Currently the package supports a flexible mixture of shifted poisson distributions. The shifted poisson augments the poisson distribution with a location parameter. More mixtures of discrete and continuous distributions are in the pipeline.

#### Installing BayesMultiMode
```{r, eval = FALSE}
# Installing from github :
# install.packages("devtools") # if devtools is not installed 
devtools::install_github("paullabonne/BayesMultiMode")
```

#### Loading the package
```{r}
library(BayesMultiMode)
```

#### Generating data
```{r}
set.seed(1)
p1 = 0.3
p2 = 1-p1
kap1 = 3
kap2 = 0
lam1 = 1
lam2 = 0.5
length_data = 70
simulated_data <- c(rpois(length_data*p1, lam1)+kap1, rpois(length_data*p2, lam2)+kap2)
```

#### Choosing either simulated or DNA data
```{r}
# Select DNA data :
data("d4z4")
y = d4z4

# Or select simulated data :
# y = simulated_data
```

#### Setting parameters for SFM MCMC estimation
```{r}
# Number of MCMC iterations 
M = 5000 

# Proportion of draws to discard as burnin
S = 0.5 

# Maximum number of mixture components 
Jmax = 6
```

#### Estimation with SFM MCMC
```{r, out.width = '70%', fig.align = "center"}
#Bayesian estimation
sfm_mcmc = sfm_mcmc_spmix(y=y,Jmax=Jmax, M=M)

#Plots of the estimation output
graphs = plots_mcmc(sfm_mcmc,S)
graphs[[3]]
graphs[[4]]
```

#### Post-processing : burn-in and discarding empty components
```{r}
post_sfmmcmc = post_sfm_mcmc(sfm_mcmc,S)
```

#### Mode inference
```{r, out.width = '70%', fig.align = "center"}
sfm_mcmc_modes = bayes_mode(post_sfmmcmc$theta_draws_slim,y)
sfm_mcmc_modes$graphs
```
