---
title: "BayesMultiMode"
output: github_document
bibliography: inst/REFERENCES.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(dpi=600)
```
<!-- badges: start -->
[![R-CMD-check](https://github.com/paullabonne/BayesMultiMode/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/paullabonne/BayesMultiMode/actions/workflows/R-CMD-check.yaml)
[![CRAN_Status_Badge](https://www.r-pkg.org/badges/version/BayesMultiMode)](https://cran.r-project.org/package=BayesMultiMode)
<!-- badges: end -->

`BayesMultiMode` is an R package for detecting and exploring multimodality using Bayesian techniques. The approach works in two stages. First, a mixture distribution is fitted on the data. The number of mixture components does not have to be specified; the size of the mixture is estimated simultaneously with the mixture weights and components through a Sparse Finite Mixture approach. Second, the resulting MCMC draws are used to calculate the number of modes and their locations. Posterior probabilities are retrieved for both of these quantities providing a powerful tool for mode inference. 

### Installing BayesMultiMode from CRAN
```{r, eval = FALSE}
install.packages("BayesMultiMode")
```

### Or installing the development version from GitHub
```{r, eval = FALSE}
# install.packages("devtools") # if devtools is not installed 
devtools::install_github("paullabonne/BayesMultiMode")
```

### Loading BayesMultiMode
```{r}
library(BayesMultiMode)
```

### Using BayesMultiMode for both MCMC estimation and mode inference
`BayesMultiMode` provides a very flexible and efficient MCMC estimation approach : it handles mixtures with unknown number of components and supports a comprehensive range of mixture distributions, both continuous and discrete.

#### Estimation
```{r, out.width = '70%', fig.align = "center"}
# retrieve acidity data
y = multimode::acidity

# estimation
bayesmix = bayes_estimation(data = y,
                            K = 5,
                            dist = "skew_normal",
                            nb_iter = 2000,
                            burnin = 1000)

# plot estimated mixture
plot(bayesmix, max_size = 200)
```

#### Mode inference
```{r, out.width = '70%', fig.align = "center"}
# mode estimation
bayesmode = bayes_mode(bayesmix)

# plot 
plot(bayesmode, max_size = 200)

# Summary 
summary(bayesmode)
```


### Using BayesMultiMode for mode inference with external MCMC output
`BayesMultiMode` also works on MCMC output generated using external software. The function `new_BayesMixture()` creates an object of class `BayesMixture` which can then be used as input in the mode inference function of `bayes_mode()`. Here is an example where the `BNPmix` package is used for MCMC estimation.
```{r, out.width = '70%', fig.align = "center"}
library(BNPmix)

y = multimode::acidity

## estimation
PY_result = PYdensity(y,
                     mcmc = list(niter = 10000, nburn = 5000),
                     output = list(out_param = TRUE))

plot(PY_result)
```

#### Transforming the output into a mcmc matrix with one column per variable
```{r, message = FALSE}
library(dplyr)

mcmc_py = list()

for (i in 1:length(PY_result$p)) {
  k = length(PY_result$p[[i]][, 1])
  
  draw = c(PY_result$p[[i]][, 1],
           PY_result$mean[[i]][, 1],
           sqrt(PY_result$sigma2[[i]][, 1]),
           i)
  
  names(draw)[1:k] = paste0("theta", 1:k)
  names(draw)[(k+1):(2*k)] = paste0("mu", 1:k)
  names(draw)[(2*k+1):(3*k)] = paste0("sigma", 1:k)
  names(draw)[3*k + 1] = "draw"
  
  mcmc_py[[i]] = draw
}

mcmc_py = bind_rows(mcmc_py)

##
pars_names = c(theta = "theta",
               mu = "mu",
               sigma = "sigma")
```

#### Creating an object of class `BayesMixture`
```{r}
py_BayesMix = new_BayesMixture(mcmc = mcmc_py,
                               data = y,
                               K = (ncol(mcmc_py)-1)/3,
                               burnin = 0, # the burnin has already been discarded
                               dist = "normal",
                               pars_names = pars_names,
                               dist_type = "continuous")
```


#### Mode inference
```{r, out.width = '70%', fig.align = "center"}
# mode estimation
bayesmode = bayes_mode(py_BayesMix)

# plot 
plot(bayesmode, max_size = 200)

# Summary 
summary(bayesmode)
```


### References
[@basturk_bayes_2021]
[@malsiner-walli_model-based_2016]
[@schaap_genome-wide_2013]
